Sexual selection and extinction: a macro-evolutionary approach

### 2018-06-27 Sexual_Selection_and_Extinction.Rmd
Initial commit

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(pander)
library(knitr)
library(tidyr)
library(mgcv)
library(diversitree)
library(repmis)
library(lme4)
library(ape)
library(geiger)
library(nlme)
library(phytools)
library(brms)
library(ggridges)
library(caper)
source("bamm_extinction/functions/check_and_fix_ultrametric.R")
source("functions/essim.R")
```

@@

Extinction rate ~ Sexual selection
                   + Range size
                   + Short temporal variability of temperature (mean BIOCLIM4)
                   + Short temporal variability of precipitation (mean BIOCLIM15)
                   + Spatial variability of temperature (PCA1)
                   + Spatial variability of precipitation (PCA2)
                   + Long-term variability of temperature (BIOCLIM1)
                   + Long-term variability of precipitation (BIOCLIM12)
                   + Variability of NPP

@@

Or perhaps OU model rather than Brownian:

```{r}
pgls.model.3 <- gls(rab.ex.rate.split ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                correlation = corMartins(1, phy = pruned.tree), 
                data = restricted.data, 
                method = "REML")
summary(pgls.model.3)
```

@@

```{r, eval=FALSE}
model_bayesian <- brm(
  rab.ex.rate.split ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res, data = restricted.data, 
  family = gaussian(), 
  cov_ranef = list(phylo = A),
                seed = 1,
                cores = 4, chains = 4, iter = 4000, #Run 4 chains in parallel for 4000 iterations (2000 are burn in)
                control = list(adapt_delta = 0.99, max_treedepth = 15))
```

```{r}
summary(model_bayesian)
```

Try looking at speciation rate using ES-Sim adapted by @Harvey_essim_2017 and found here https://github.com/mgharvey/ES-sim

```{r}
Essim.frame <- restricted.data$SDi
names(Essim.frame) <- restricted.data$TipLabel

essim.results <- essim(pruned.tree, Essim.frame, nsim = 1000, return.es = TRUE)
```

We can then add the essim.results and run a similar PGLS model
```{r}
restricted.data$ESsim <- essim.results[["es"]]

restricted.data %>% ggplot(aes(x=BIO12.LGM.Diff, y = (ESsim))) + 
  geom_point()+
  geom_smooth(method = "lm")
```

 Once we have our EsSim results we can use them in a linear model to assess the relationship between speciation and our predictors

```{r}
model.2 <- lm(ESsim ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                data = restricted.data)
summary(model.2)
```

Have a look at how these would look if we obtained model predictions for each data point.

```{r, fig.width=10, fig.height=5}
predict.frame <- cbind(predict(model.2),restricted.data)
predict.frame %>% ggplot(aes(y = `predict(model.2)`, x = SDi)) + 
  geom_point(aes(color = BLFamilyLatin))+
  scale_color_hue(h = c(0,100), l = 50) +
  geom_smooth(color = 'black')+
  theme_minimal()
```

It looks like some of the higher SDi's have slightly higher speciation rates. Perhaps we can look at those obviously dimorphic groups as category of its own.

```{r}
restricted.data %>% ggplot(aes(SDi))+
  geom_histogram(binwidth = 0.015)+
  theme_minimal()
```

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(pander)
library(knitr)
library(tidyr)
library(mgcv)
library(diversitree)
library(repmis)
library(lme4)
library(ape)
library(geiger)
library(nlme)
library(phytools)
library(brms)
source("bamm_extinction/functions/check_and_fix_ultrametric.R")
```

@@

Extinction rate ~ Sexual selection
                   + Range size
                   + Short temporal variability of temperature (mean BIO4)
                   + Short temporal variability of precipitation (mean BIO15)
                   + Spatial variability of temperature
                   + Spatial variability of precipitation
                   + Long-term variability of temperature
                   + Long-term variability of temperature
                   + Variability of NPP

@@

How about a PGLS model that uses the IUCN rankings as a categorical response:

```{r}
pgls.model.2 <- gls(X2010.IUCN.Red.List.category ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res,
                correlation = corBrownian(phy = pruned.tree), 
                data = restricted.data, 
                method = "REML")
summary(pgls.model.2)
```

@@

```{r}
model_simple <- brm(
  rab.ex.rate.split ~ SDi 
                + range.size.m2 
                + bioclim4
                + bioclim15
                + residuals.PC1
                + residuals.PC2
                + BIO1.LGM.Diff
                + BIO12.LGM.Diff
                + NPP.res, data = restricted.data, 
  family = gaussian(), 
  cov_ranef = list(phylo = A),
                seed = 1,
                cores = 4, chains = 4, iter = 4000, #Run 4 chains in parallel for 4000 iterations (2000 are burn in)
                control = list(adapt_delta = 0.99, max_treedepth = 15))
```

```{r}
summary(model_simple)
```
